### 2019-05-28 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Done for merging Xtof's comments on Ch4

The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
and the network is less and less stationary.
% and the stochastic hypothesis is more and more violated.
To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions. Static devices form an interfering traffic, which could be generated by devices using other standards as well.
Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel. So our first hypothesis is actually not constraining.
We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
This second approach is ideal, but not realistic in decentralized IoT networks, and it is just used here as a reference.
Then, these three reference scenarios allow to evaluate the performance of \UCB{} and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.

@@

\paragraph{One gateway and many devices}
%
We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of devices all sends uplink packets to a (unique) network gateway.
Messages can be sent in a fixed number $K\geq2$ of wireless channels,
that are assumed to be orthogonal, that is, a message in channel $k$ will not cause collision to a message in another channel $j \neq i$.
% (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
%
% Note that this choice of number of channels and of frequency bands is for instance the one of the current LoRa standard in Europe, but of course the precise choice does not really matter, and future standards can choose different parameters.

@@

The devices can transmit their packets in one of the $K$ channels. In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
This is realistic in IoT networks such as networks based on LoRa.
% (of $1\;\mathrm{s}$ in our implementation).
In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the device by letting only one gateway sends back an acknowledgement.
The natural choice is for the network to send this acknowledgement by the gateway which received the uplink message.
For simplicity but without loss of generality, we consider a network with only one gateway,
but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming uplink messages from the IoT devices being paired in the network.
%

@@

\paragraph{Static vs dynamic devices}
%
We make the following assumptions on the network.
We assume that there are two types of end-devices in the network:
\begin{itemize}
    \item
    \emph{Static} end-devices are identical, and each of them uses only \emph{one} channel to communicate with the gateway, with no loss of generality.
    %  if devices random access the channels.
    Their choice is assumed to be fixed in time (\ie, stationary). The traffic generated by these devices is considered as an interfering traffic for other devices.
    \item
    \emph{Dynamic} (or \emph{smart}) end-devices can use all the available channels, by quickly changing their communication channel at any time slot, following a Machine Learning policy.
    For that end, they use their history of past communication successes or failures they experienced in each channel, in order to learn about channel availability.
    We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bits IEEE float number for each of the $K$ channels).
    Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, to improve their successful communication rate automatically.
\end{itemize}

We further assume that there are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices.
Furthermore, in channel $k \in \llbracket 1; K \rrbracket$ there are $0 \leq S_k \leq S$ static devices (so $S = \sum_{k=1}^{K} S_k$).

@@

We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in \llbracket 1; K \rrbracket$).
This reflects a naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.

>>>>

The aim of this section is to assess the potential gain of learning algorithms in IoT scenarios, even when the number of intelligent devices in the network increases,
and the network is less and less stationary.
% and the stochastic hypothesis is more and more violated.
To do that, we suppose an IoT network made of two types of devices: \emph{static devices} that use only one channel (fixed in time), and \emph{dynamic devices} that can choose the channel for each of their transmissions. Static devices form an interfering traffic, which could be generated by devices using other standards as well.
Note that instead of assuming that each static device uses a fixed channel, we could also assume a looser hypothesis: if each static device uses a fixed sub-set of the $K$ channels, and a purely uniform random access in its set of considered channels, then \emph{in average} the observed occupancy of the $K$ channels can be modeled as if it were occupied by (more) static devices using only one channel. So our first hypothesis is actually not constraining.
We first evaluate the probability of collision if dynamic devices randomly select channels (that is, a naive approach), and if a centralized controller would optimally distribute all of them in the channels at the beginning of the scenario.
This second approach is ideal, but not realistic in decentralized IoT networks, and it is just used here as a reference.
Then, these three reference scenarios allow to evaluate the performance of bandit algorithms, such as \UCB{} and TS, in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
We show that these algorithms have near-optimal performance, even when the proportion of end-devices increases and the interfering traffic from other devices becomes less and less stochastic.

@@

\paragraph{One gateway and many devices}
%
We consider the system model presented in Figure~\ref{fig:41:system_model1}, where a set of devices all sends uplink packets to a (unique) network gateway.
Messages can be sent in a fixed number $K\geq2$ of wireless channels,
that are assumed to be orthogonal, that is, a message in channel $k$ will not cause collision to a message in another channel $j \neq i$.
% (\eg, $K=4$ or $K=16$ for instance), which are centered indifferently around the $433.5\;\mathrm{MHz}$ or the $877\;\mathrm{MHz}$ ISM bands (as it is chosen for the implementation presented in Section~\ref{sec:4:gnuradio}).
The communication between IoT devices and this gateway is done through a simple pure acknowledged ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
We mean here by ALOHA-based that devices do not use sensing, and we do not consider retransmissions of packets (see Section~\ref{sec:4:retransmissions} below where we remove this last hypothesis).
%
% Note that this choice of number of channels and of frequency bands is for instance the one of the current LoRa standard in Europe, but of course the precise choice does not really matter, and future standards can choose different parameters.

@@

The devices can transmit their packets in one of the $K$ channels. In the case where the gateway --of the corresponding IoT network-- receives an uplink message in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay.
This is realistic in IoT networks such as networks based on LoRaWAN.
% (of $1\;\mathrm{s}$ in our implementation).
In a more general setting, if there were more than one gateway in the same IoT network being considered, the network would reply to the device by letting only one gateway sends back an acknowledgement.
The natural choice is for the network to send this acknowledgement by the gateway which received the uplink message.
For simplicity but without loss of generality, we consider a network with only one gateway,
but still it is important to observe the distinction between the gateway, which is simply a RF node, and the IoT network in charge of receiving, handling, and replying to the incoming uplink messages from the IoT devices being paired in the network.
%

@@

\paragraph{Static vs dynamic devices}
%
We make the following assumptions on the network.
We assume that there are two types of end-devices in the network:
\begin{itemize}
    \item
    \emph{Static} end-devices are identical, and each of them uses only \emph{one} channel to communicate with the gateway, with no loss of generality.
    %  if devices random access the channels.
    Their choice is assumed to be fixed in time (\ie, stationary). The traffic generated by these devices is considered as an interfering traffic for other devices.
    \item
    \emph{Dynamic} (or \emph{smart}) end-devices can use all the available channels, by quickly changing their communication channel at any time slot, following a Machine Learning policy.
    For that end, they use their history of past communication successes or failures they experienced in each channel, to learn about channel availability.
    We also assume that the dynamic end-devices can run simple embedded decision making algorithms, and have limited but reasonable computing as well as storage capacities.
    Note that of course we assume a limited storage capacity, so no device stores the full history of communication successes and failures, but they only store average rates (which can be stored using one $32$-bit IEEE 754 float number for each of the $K$ channels).
    Our goal is to propose a learning algorithm than can be implemented by each dynamic device, in a decentralized and independent manner, to improve their successful communication rate automatically.
\end{itemize}

We further assume that there are $K \geq 2$ channels, $D \geq 0$ dynamic end-devices, and $S \geq 0$ static devices
with $0 \leq S_k \leq S$ static devices in channel $k \in [K]$ (so $S = \sum_{k=1}^{K} S_k$).

@@

We derive here the probability of having a successful transmission, for a dynamic device, in the case where all the dynamic devices make a purely random channel selection (\ie, uniform on $i \in [K]$).
This reflects a naive policy that could be implemented by all the dynamic devices, and it provides a reference scenario to compare against.

### 2019-05-28 2-Chapters/4-Chapter/chapter4.tex
Small proofreading of notations of Chapter6
My thesis is so long, and so stupid
Done for merging Xtof's comments on Ch4

% ----------------------------------------------------------------------
\chapter{Improving Spectrum Usage of IoT Networks with Selfish MAB Learning}
\label{chapter:4}
\minitoc
% ----------------------------------------------------------------------

@@

\newpage
% Write miniTOC just after the title
\graphicspath{{2-Chapters/4-Chapter/Images/}}

@@

Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
As end-devices battery life is a key constraint of IoT networks,
and as IoT networks are decentralized (\ie, the devices initiate transmissions),
this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
%
In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
Kaufmann12Thompson}, for the first model. For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.

@@

% \newpage  % WARNING ?
% ----------------------------------------------------------------------------
\section{Conclusion}
\label{sec:4:conclusion}
% ----------------------------------------------------------------------

>>>>

% ----------------------------------------------------------------------
\chapter{Improving Spectrum Usage of IoT Networks with Selfish MAB Learning}
\label{chapter:4}
\minitoc
% Write miniTOC just after the title
\graphicspath{{2-Chapters/4-Chapter/Images/}}
% ----------------------------------------------------------------------

\TODOL{Intégrer les commentaires de Christophe, le plus rapidement possible !}

@@

\newpage

@@

Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
As end-devices battery life is a key constraint of IoT networks,
and as IoT networks are decentralized (\ie, the devices initiate transmissions),
this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
%
In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, that could be used in combination with a time-frequency slotted ALOHA-based protocol.
Even without changing anything of the level of IoT standards, our proposal is just an add-on capability that can be used on a unit-per-unit basis.
We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
Kaufmann12Thompson}, for the first model. For the demonstration as well as for the second model, without loss of generality, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored to solve the problem of learning in order to retransmit efficiently.

@@

\newpage  % WARNING ?
% ----------------------------------------------------------------------------
\section{Conclusion}
\label{sec:4:conclusion}
% ----------------------------------------------------------------------
